---
title: "HW1_answer_complete"
output: html_document
---

```{r}
library(dplyr)
library(lubridate)
library(tidyr)
library(stringr)

```

#Reading Farmers Market Data and Wine data
```{r}
fm_df <- read.csv('fm.csv')
wine_df <- read.csv('wine_data.csv')
```

#1.1  Compute the number of farmers market in the United States
```{r}
num_markets <- length(unique(fm_df$FMID))
print(paste0("Number of farmers markets in United States: ", num_markets))
```

#1.2  Compute the number of farmers market in Massachusetts
```{r}
num_markets_ma <- dim(fm_df[fm_df$State == "Massachusetts", ])[1]
print(paste0("Number of farmers markets in Massachusetts: ", num_markets_ma))

```

#1.3 Count the number of farmers market that offer coffee in the state of California
```{r}
coffee_count <- fm_df %>%
                filter(State == "California", Coffee == "Y")
print(paste0("Farmers markets with Coffee in California: ", dim(coffee_count)[1]))
```

#1.4 Separate Season1Date into Startdate and Enddate, Calculate SeasonDuration (In days)
```{r}
#Split Season1Date into Startdate and Enddate fields
x <- fm_df %>% 
  separate(Season1Date, into = c("Startdate", "Enddate"), sep = " to ")

fm_df$Startdate <- x$Startdate
fm_df$Enddate <- x$Enddate

#Parse dates : convert all month names to start of month eg "May" to 0000-05-01 UTC
fm_df$Startdate <- parse_date_time(fm_df$Startdate, c("b", "mdy"))
fm_df$Enddate <- parse_date_time(fm_df$Enddate, c("b", "mdy"))

#Create a duration object to calculate season duration, will work for most cases, handle extra cases below
duration_vec <- fm_df$Startdate %--% fm_df$Enddate
season_duration <- as.duration(duration_vec) / ddays(1)
fm_df$SeasonDuration <- season_duration

#-------------------------------------------------------------------------------------------------------------
# Handling season duration (negative) for cases like 
#   1. November - April (gets parsed to 0000-11-01 UTC to 0000-04-01 UTC)
#   2. 2012-06-11 to October etc 
#-------------------------------------------------------------------------------------------------------------
  
#Filter rows with negative duration, ignore FMID 1011959 as it has invalid data entry for Season1Date = 10/01/2016 to 05/07/2016 (m/d/y)
neg_duration <- fm_df %>%
                filter(SeasonDuration < 0)  

#Adding one year to End date year for cases like November to April (lubridate assigns year 0 if only month names are present)
year(neg_duration[month(neg_duration$Startdate) > month(neg_duration$Enddate) & (year(neg_duration$Startdate) == 0) & (year(neg_duration$Enddate) == 0), ]$Enddate) <- 1

#Handling year for cases like 2012-06-11 to October (Startdate year is non-zero, Enddate year will be 0),
#Assign Startdate year to Enddate
neg_duration_end_month <- neg_duration[month(neg_duration$Startdate) < month(neg_duration$Enddate) & (year(neg_duration$Startdate) != 0) & (year(neg_duration$Enddate) == 0), ]
year(neg_duration_end_month$Enddate) <- year(neg_duration_end_month$Startdate)

#Replace Enddates to include above calculated end dates, create one final dataframe
neg_duration$Enddate <- replace(neg_duration$Enddate, as.numeric(rownames(neg_duration_end_month)), neg_duration_end_month$Enddate)

#Correct the season duration for these entries, same as above
duration_vec <- neg_duration$Startdate %--% neg_duration$Enddate
season_duration <- as.duration(duration_vec) / ddays(1)
neg_duration$SeasonDuration <- season_duration

fm_df$SeasonDuration[match(neg_duration$FMID, fm_df$FMID)] <- neg_duration$SeasonDuration

print(fm_df)
```

#2.1 Calculate 20 year old Wine 
```{r}
ind <- str_which(wine_df$designation, regex("20.yr|20.year", ignore_case = TRUE))
wine_number <- dim(wine_df[ind,])[1]
print(paste0("Number of 20 year old wines: ", wine_number))

```

#2.2 Average wine prices by country
```{r}
wine_agg <- wine_df %>% 
        group_by(country) %>%
        summarise(mean_price = mean(price, na.rm=TRUE))
print("Average wine prices by country:")
print(wine_agg)

#Same aggregation using for loop:
#---------------------------------------------------

countries <- unique(wine_df$country)
agg_df <- data.frame(country = character(),
                     mean_price = double())

#filter by country and calculate mean
for (c in countries){
  temp_df <- wine_df[wine_df$country == c, ]
  mean_value <- mean(temp_df$price, na.rm = T)
  result_df <- data.frame(country = c(c),
                          mean_price = c(mean_value)
  )
  agg_df <- rbind(agg_df, result_df) #Bind new entries
  agg_df <- agg_df %>%    #Sort by country
    arrange(country)
}
print(agg_df)
#-----------------------------------------------------
```
